---
title: "Final Figures Report"
author: "Madison Volpe"
date: "5/11/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(tidyverse)
library(ggplot2)
library(sf) # for shapefiles
library(maptools)
library(rgdal)
library(rgeos)
library(reshape2)
library(scales)
library(kableExtra)
```

```{r}
# read in drug arrests 
drugs <- read.csv("./data/si_drugs_mod_arrests.csv")

# read in zip shapefile 
zip   <-readOGR(dsn = "./shapefiles/ZipCode")

# convert zip to sf 
zip <- st_as_sf(zip)
zip <- zip[,c(1,4)]

# filter for SI zip codes only 
zip <- filter(zip, zcta %in% c(10302, 10303, 10310, 10306, 10307, 10308, 10309, 
                               10312, 10301, 10304, 10305, 10314, 10311, 10313))

colnames(zip) <- c("zip", "geometry")

# transfom zip to 4326 
zip <- st_transform(zip, 4326)

# get points from drugs dataset 
points <- data.frame(x = drugs$longitude,  y = drugs$latitude, id = 1:nrow(drugs), race = drugs$perp_race) 
points <- st_as_sf(points, coords = c("x", "y"), crs = 4326)

# Intersection between polygon and points -
intersection <- st_intersection(x = zip, y = points)
intersection <- as.data.frame(intersection)
intersection <- intersection[1:3]
intersection <- as.data.frame(intersection)
intersection$race <- as.character(intersection$race)

# Counts 
race.results <- intersection %>%
                  select(zip,race) %>%
                  group_by(zip,race) %>%
                  count()

names(race.results)[3] <- "Amount"

race.results <- tidyr::spread(data= race.results, key = race, value = Amount)
```

```{r}
make.zero <- function(x){
  x[is.na(x)] <- 0
  return(x)
}

make.numeric <- function(x){
  x <- as.numeric(as.character(x))
  return(x)
}

race.results <- data.frame(apply(race.results, 2, make.zero))
race.results <- data.frame(apply(race.results, 2, make.numeric))

race.results <- race.results %>%
  mutate(OTHER = AMER.IND + AMERICAN.INDIAN.ALASKAN.NATIVE + ASIAN...PACIFIC.ISLANDER +
           ASIAN.PAC.ISL+ UNKNOWN,
         TOTAL = WHITE + BLACK + HISPANIC + OTHER) %>%
  select(zip, WHITE, BLACK, HISPANIC, OTHER, TOTAL)

zip.results <- read.csv("./data/ModelDatasets/zipcodeModel3.csv")

final <- left_join(race.results, zip.results, by = 'zip')

final <- final %>%
  select(zip, WHITE, BLACK, HISPANIC, OTHER, TOTAL, Total.Population.Over18, proportionWhite, proportionNonWhite)%>%
  mutate(MajorityNonWhite = ifelse(proportionNonWhite > .5, "Yes", "No")) %>%
  mutate(DrugArrestRate = (TOTAL/Total.Population.Over18)*100) %>%
  select(zip, WHITE, BLACK, HISPANIC, OTHER, TOTAL, MajorityNonWhite, DrugArrestRate) %>%
  arrange(desc(DrugArrestRate))
```

```{r}
kableExtra::kable(final, format = 'latex', booktabs = T, caption = "Drug Arrests by Zip Code",
                  col.names = c("Zip","White", "Black", "Hispanic", "Other", "Total", "NonWhite Majority", "Arrest Rate")) %>%
    kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F) 
  
```

\begin{table}[!htbp] \centering 
  \caption{Poisson Regression Results} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & Drug Arrests (2013-2018) & Drug Arrests (2014-2017) & Drug Arrests (2013-2018) \\ 
 & Aggregated &  & Aggregated \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 EMS Calls 2013-2018 & 0.0001 &  & 0.0005 \\ 
  & (0.0001) &  & (0.0003) \\ 
  EMS Calls 2014-2017 &  & 0.001$^{**}$ &  \\ 
  &  & (0.0003) &  \\ 
  Proportion Non-White & 2.070$^{***}$ & 2.126$^{***}$ & 1.320$^{*}$ \\ 
  & (0.467) & (0.297) & (0.563) \\ 
  Naloxone Saves (2016-2018) &  &  & $-$0.007 \\ 
  &  &  & (0.006) \\ 
  Drug Overdose Deaths (2016-2018) &  &  & $-$0.005 \\ 
  &  &  & (0.012) \\ 
  Constant & $-$4.369$^{***}$ & $-$6.073$^{***}$ & $-$3.680$^{***}$ \\ 
  & (0.251) & (0.160) & (0.447) \\ 
 \hline \\[-1.8ex] 
Observations & 12 & 48 & 12 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 
